---
title: "DATA607_HW2"
author: "Dilip Ganesan"
date: "02/09/2017"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 
if (!require("htmlTable")) install.packages('htmlTable')
library(htmlTable)
if (!require("ggplot2")) install.packages('ggplot2')
library(ggplot2)
if (!require("DBI")) install.packages('DBI')
library(DBI)
if (!require("RMySQL")) install.packages('RMySQL')
library(RMySQL)

# Some customization.  You can alter or delete as desired (if you know what you are doing).
trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```

## DATA 607 Home Work 2 - R and SQL.


### Introduction.

In this Home Work, I have tried to connect to MYSQL database, get records from table and have created data.frames

***SQL Tables:*** 
Have created three tables for this home work namely REVIEWER, MOVIE_NAMES and REIVIEW_MOVIE_RATINGS 
Wanted to try Relational database model, so created three tables rather than one.

***Approach*** consist of manipulating the data set in two ways.

*1. Using the SQL query joins *

*2. Using the R programming functions to achieve resulset data.frame *


### Step 1: SQL Connection

```{r}
#Connecting to MySQL database using dbConnect. Password is not masked for home work purpose.
mydb = dbConnect(MySQL(), user='root', password='mysql@123', dbname='DATA_607', host='localhost')
dbListTables(mydb)
```

### Step 2: Fetching records from tables.

```{r}
# Now trying to get the 3 table data as individual data.frames.
reviewer <- dbGetQuery(mydb, "select * from reviewer")

movie_names <- dbGetQuery(mydb, "select * from movie_names")

ratings <- dbGetQuery(mydb, "select * from review_movie_rating")
```

### Step 3: Checking how data got populated in data frames..

```{r}
head(reviewer)
head(movie_names)
head(ratings)
```

### Step 4: Using the SQL query to attain resultant data frame.
```{r}
review_movie_rating= dbGetQuery(mydb, "SELECT A.REVIEWER, B.MOVIE_NAMES, C.RATINGS FROM REVIEWER A, MOVIE_NAMES B, REVIEW_MOVIE_RATING C WHERE C.REVIEWER_ID = A.REVIEWER_ID AND C.MOVIE_ID = B.MOVIE_ID ORDER BY A.REVIEWER;")

dim(review_movie_rating)
# So the resultset contains 36 rows and 3 variables

# Cleaning out outstanding database connection.
dbDisconnect(mydb)
```

### Step 5: Printing Resultant Data Frame.
```{r}
htmlTable(review_movie_rating, caption = '2017 SPRING CUNY MSDA CLASS MOVIE REVIEW RATINGS')
```

### Step 6: Using R Programming functions to create Resultant Data Frame.
```{r}
# Used the merge function to merge the data set.

dt<-merge(reviewer,merge(movie_names,ratings,by="movie_id"), by="reviewer_id")
finaldata= subset(dt,select = c(2,4,5))
```

### Step 7: Using ggPlot to create a small plot..
```{r}
# Plot shows the Mean Ratings across the movies.
ggplot(review_movie_rating) + geom_bar(aes(review_movie_rating$MOVIE_NAMES, review_movie_rating$RATINGS, fill=review_movie_rating$MOVIE_NAMES), stat="summary", fun.y="mean") + labs(x="Movies", y="Mean Review", title="Mean Ratings of Top American Movies") + theme(axis.text.x=element_blank())


ggplot(review_movie_rating) + geom_bar(aes(review_movie_rating$REVIEWER , review_movie_rating$RATINGS, fill=review_movie_rating$REVIEWER), stat="summary", fun.y="mean") + labs(x="Movies", y="Mean Review", title="Mean Ratings of Top American Movies") + theme(axis.text.x=element_blank())
```



  * File creation date: `r Sys.Date()`
  * `r R.version.string`
  * R version (short form): `r getRversion()`
  * `mosaic` package version: `r packageVersion("mosaic")`
  * Additional session information
  

  