---
title: "Data 607 Project 1"
author: "Dilip Ganesan"
date: "02/24/2016"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}
require(mosaic)   # Load additional packages here 

if (!require("stringr")) install.packages('stringr')
if (!require("tidyr")) install.packages('tidyr')
library(tidyr)
library(stringr)

options(warn=-1)

# Some customization.  You can alter or delete as desired (if you know what you are doing).
trellis.par.set(theme=theme.mosaic()) # change default color scheme for lattice
knitr::opts_chunk$set(
  tidy=FALSE,     # display code as typed
  size="small")   # slightly smaller font for code
```


### Project 1 : Chess Data Transformation.

### Introduction.

This Project consist of text file which contains the rankings and players information of a Chess Tournament.
The data available should be processed and displayed in the format specified in the project.

***Approach***

The entire data is divided in to two data frames.

*1. With First data frame contains the player related information.*

*2. Second data frame contains the ranking related information.*

*3. Final result is displayed using the two database as two relational database tables.*

***Step 1: Loading the text file and cleansing unwanted elements.***

```{r}
datafile=readLines("tournamentinfo.txt")
#Cleaning the unwanted "-----"
datafile=str_replace_all(datafile,"-----------------------------------------------------------------------------------------","")
```

***Step 2: Creating a dataframe and loading that file to data frame, cleansing empty rows in dataset.***


```{r}
#Creating a dataframe and loading that file to data frame.
df=data.frame()
df=data.frame(datafile)

#Cleaning the unwanted empty rows for clear data set.
df[df==""] <- NA
df = df %>% na.omit()


```

***Step 3: Creating 1st Data Frame with Player Information.***

```{r}
# 1st Data.Frame.
df5=data.frame()
toDelete <- seq(1, nrow(df), 2)
df5<-data.frame(df[ toDelete ,])
colnames(df5)

df5=separate(data = df5, col = df.toDelete..., into = c('Pair','Player Name','Total','R1','R2','R3','R4','R5','R6','R7' ), sep = "\\|")
df5 <- df5[-c(1), ]


```

***Step 4: Creating 2nd Data Frame with Player Ranking Information***

```{r}
# 2nd Data.Frame.
toDelete2 <- seq(2, nrow(df), 2)
df2<-data.frame(df[ toDelete2 ,])
colnames(df2)
df3=data.frame()
df3=separate(data = df2, col = df.toDelete2..., into = c('State','USCF','Pts','1','2','3','4','5','6','7' ), sep = "\\|")
df3<-df3[-c(1),]


```

***Step 4: This is a function to calculate Average of Opponents Pre Ratings.***


```{r}
# This method is used to calculate the Average Pre Rating for Opponents.

avgrat<-as.vector(c())
for (i in 1:nrow(df5)){
  oc<-str_extract_all(df5[i,],"\\b\\d{1,}")
  oc<-as.matrix(oc[-1:-3])
  oc<-as.matrix(oc[lapply(oc,length)>0])
  total=0
  k=0
  j=0
  for(row in 1:nrow(oc)){
    k=oc[row]
    presub2<- str_extract(df3[as.numeric(k),]$USCF, ': *\\d{2,}')
    presub2<- str_extract(presub2, '\\d{2,}')
    total=total+as.numeric(presub2)
    j=j+1
  }
  avg=total/j
  print(round(avg))
  avgrat[length(avgrat)+1]<-round(avg)
}


```


***Step 5: Final Data Cleaning and Resultant Data frame.***


```{r}
# Final Data Cleansing.
prerating <- str_extract(df3$USCF, ': *\\d{2,}')

#Resultant Data Frame.
df6<-data.frame()
df6=data.frame(PlayerName=df5$`Player Name` ,State=df3$State ,TotalPoints=df5$Total , 
               PreRating=as.integer(str_extract(prerating, '\\d{2,}')), AvgOppPreRating=avgrat)

write.csv(df6, "ChessRankingData.csv", row.names=FALSE)

```

***References***
StackOverflow for figuring the Seperate method to split a column in data frame.