---
title: "DATA607_HW9"
author: "Dilip Ganesan"
date: "04/29/2017"
output:
  pdf_document:
    fig_height: 3
    fig_width: 5
  html_document:
    fig_height: 3
    fig_width: 5
  word_document:
    fig_height: 3
    fig_width: 5
---

```{r, setup, include=FALSE}

if (!require("DBI")) install.packages('DBI')
library(DBI)
if (!require("RMySQL")) install.packages('RMySQL')
library(RMySQL)
if(!require("mongolite"))install.packages("mongolite")
library(mongolite)


```

## DATA 607 Home Work 9 - R/MySQL/MongoDB.


### Introduction.

In this Home Work, I have tried to connect to MYSQL database, get records from table and have created data.frames and migtrated the data to MongoDB.

***SQL Tables:*** 
Have created three tables for this home work namely REVIEWER, MOVIE_NAMES and REIVIEW_MOVIE_RATINGS 

***Mongo***

In MongoDB, everything is just a Collection. So all the individual fields will get moved as a single collection.



### Step 1: SQL Connection

```{r}
#Connecting to MySQL database using dbConnect. Password is not masked for home work purpose.
mydb = dbConnect(MySQL(), user='root', password='mysql@123', dbname='DATA607', host='localhost')
dbListTables(mydb)
```

### Step 2: Fetching records from tables.

```{r}
# Now trying to get the 3 table data as individual data.frames.
reviewer <- dbGetQuery(mydb, "select * from reviewer")

movie_names <- dbGetQuery(mydb, "select * from movie_names")

ratings <- dbGetQuery(mydb, "select * from review_movie_rating")
```

### Step 3: Checking how data got populated in data frames..

```{r}
head(reviewer)
head(movie_names)
head(ratings)
```

### Step 4: Connecting to MongoDB using MongoLite.
```{r}
##if (!require("RMongo")) install.packages('RMongo')
 ##library(RMongo)

##mongo = mongoDbConnect("test", "127.0.0.1", 27017)

##b=mongo.bson.from.df(reviewer)

##output <- dbInsertDocument(mongo, "reviewer", reviewer)
##output <- dbGetQuery(mongo, 'test_data', '{"foo": "bar"}')
##print(output)

# Connectiong to MongoDB using MongoLite pacakge
con=mongo(collection = "test", db = "test", url = "mongodb://localhost",
verbose = FALSE, options = ssl_options())



```

### Step 5: Migrating Data to Mongo DB.
```{r}
# Inserting in to MongoDB
con$insert(reviewer)
con$insert(movie_names)
con$insert(ratings)

# Getting the count of rows
con$count()

#Find Query to select all fields
alldata=con$find('{}')
knitr::kable(alldata)

#Remove all data from MongoDB
con$remove('{}')

#Drop the collection
con$drop()
```




